---
title: "Homework 4"
author: "Wenfeng Zhang"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: yes
    toc: true
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Homework 4 vignette}
-->

## 5.8.2

By CASL equation 2.73, 
$$cond(A)=\frac{\sigma_{max}}{\sigma_{min}}$$
where $\sigma_{max}$ is the largest singular value and $\sigma_{min}$ is the smallest singular value, of matrix $A$  

Set
```{r}
X <- matrix(c(1, 2, 2, 1), 2, 2)
p <- c(0.5, 0.000001)
X
p
```

Then the condition number of $X'X$ is
```{r}
svals <- svd(t(X)%*%X)[["d"]]
max(svals)/min(svals)
```
The logistic variation of Hessian matrix is 
$$X'diag(p(1-p))X$$
The condition number of this Hessian matrix is 
```{r}
H = t(X)%*%diag(p*(1-p))%*%X
svals2 <- svd(H)[["d"]]
max(svals2)/min(svals2)
```

The condition number of logistic variation Hessian matrix $X'diag(p(1-p))X$ is much larger than that of $X'X$; the linear Hessian is well conditioned but the logistic variation is not.


##5.8.4

```{r}
irwls_glm <- function(X, y, family, maxit=25, tol=1e-10, lambda){
  beta <- rep(0,ncol(X))
  for(j in 1:maxit){
    b_old <- beta
    eta   <- X %*% beta
    mu    <- family$linkinv(eta)
    mu_p  <- family$mu.eta(eta)
    z     <- eta + (y - mu) / mu_p
    W     <- as.numeric(mu_p^2 / family$variance(mu))
    XtX   <- crossprod(X, diag(W) %*% X)
    Xtz   <- crossprod(X, W * z)
    # add l2 norm penalty -------------------------
    V <- XtX + diag(lambda, dim(XtX)[1])
    beta  <- solve(V, Xtz)
    # ---------------------------------------------
    if(sqrt(crossprod(beta - b_old)) < tol) break
  }
  beta
}
```

##Problem 3
Please see sparse_matrix.r in directory R
